Iterative Monomorphisation

### 2024-03-11 iterative_monomorphisation.tex
some rewritting and typesetting for the pseudocode

\textbf{TODO: Mention that all sets referenced in the algorithm are finite.}

@@

This algorithm relies on ``primitives'' whose implementation depend on the specifics of the grammar and logic used. They will therefore not be expanded upon. Functions computing the following are assumed to be available: 
\begin{enumerate}
   \item[\labelitemi] extracting the initial type argument maps from the input problem
   \item[\labelitemi] gathering type variables from a type tuple
   \item[\labelitemi] type matching
   \item[\labelitemi] returning the type variables instantiated by a substitution
   \item[\labelitemi] testing the compatibility of two substitutions
   \item[\labelitemi] substitution composition
   \item[\labelitemi] substitution application to formulae
\end{enumerate}

@@

\SetKwFunction{GenFormulae}{generate\_mono\_formulae}
\begin{figure}
\begin{quote}
\begin{algorithm}[H]
\Fn(){\(\GenFormulae(\Phi, S)\)}{
   \SetKw{And}{and}\SetKw{True}{true}\SetKw{False}{false}\SetKw{To}{to}

@@

   \ForEach{\(\varphi\in\Phi\) \textbf{such that} $\varphi$ is non-monomorphic}{
      \ForEach{\(\sigma\in \MonoSubst{\AllSubst, \(\TyVars{\(\varphi\)}\), \(\emptyset\), \(\{\}\)}\)}{
         \textcolor{ourblueviolet}{
         \eIf{\(|\NewFormulae|<\Limit\)}{
            \textcolor{black}{
            \(\NewFormulae\leftarrow\NewFormulae\cup \{\sigma(\varphi)\}\)\;}
         }
         {\Return \NewFormulae}
      }}
   }

@@

   \Return \NewFormulae
}
\end{algorithm}
\end{quote}
\caption{Pseudocode for monomorphic formula generation}
\label{genformulae}
\end{figure}

The iterative monomorphisation algorithm (\ref{iter_mono}), as implemented in Zipperposition, has two main phases each consisting of a separate loop. The first loop applies a \emph{monomorphisation step} to each formula (\ref{mono_step}) for each iteration of the outermost for loop until the user-set limit it reached. This bound is the only limit which is necessary for the algorithm to terminate. At the end of each of these iterations, the old and new type argument maps are updated with newly generated types.
%Updating the maps progressively over the course of an iteration would be feasible but would result in an uneven treatment of formulae. Type arguments in a formula processed later could be matched against monomorphic type arguments generated at an earlier time in the same iteration. As a result, the later a formula is processed the more new type arguments it could generate.

Once the first phase is completed, the substitutions used to create new type arguments are combined to form \emph{monomorphising substitutions}. These substitutions are subsequently applied to the input problem's polymorphic formulae. The possible number of new (monomorphic) formulae resulting from this process is high but can be limited by setting a maximal number of new formulae. In the Zipperposition implementation, this bound is one of the more important ones as the performance of the E prover can be significantly affected by the number of clauses it is given.

\SetKwData{MonoCap}{mono\_cap}\SetKwData{MonoMult}{mono\_mult}\SetKwData{MonoFloor}{mono\_floor}
\SetKwData{PolyCap}{poly\_cap}\SetKwData{PolyMult}{poly\_mult}\SetKwData{PolyFloor}{poly\_floor}
\begin{figure}
\begin{quote}
\begin{algorithm}[H]

   %\LinesNumbered
   \SetKwFunction{SubstGen}{matches}\SetKwFunction{TyVars}{type\_variables}
   \SetKwFunction{FMonoStep}{formula\_mono\_step}

   \Fn(){\FMonoStep{\(\varphi, M_{\text{old}}, M_{\text{new}}, N_{\text{old}}(\varphi), N_{\text{new}}(\varphi)\)}}{
   \SetKwData{TypeVars}{type\_variables}
   \SetKwData{MaxMono}{max\_mono\_args}
   \SetKwData{MaxPoly}{max\_poly\_args}
   \SetKwData{UsedSubst}{\(S\)}

@@

   \SetKwFunction{Max}{max}\SetKwFunction{Min}{min}
   \SetKwFunction{Domain}{subst\_domain}\SetKwFunction{Len}{len}

@@

   \TypeVars \(\leftarrow\) \TyVars{\(\varphi\)}\;

   \textcolor{ourblueviolet}{
   \MaxMono \(\leftarrow\) \Max{\Min{\MonoFloor, \(|M_{\text{old}}\cup M_{\text{new}}| \cdot \MonoMult\)}, \MonoCap}\;
   \MaxPoly \(\leftarrow\) \Max{\Min{\PolyFloor, \(|N_{\text{old}}(\varphi)\cup N_{\text{new}}(\varphi)| \cdot \PolyMult\)}, \PolyCap}\;
   }

@@

   \ForEach{\(\sigma\in {}\)\SubstGen{\(M_{\text{new}}\), \(N_{\text{new}}(\varphi)\)}
\(\cup\) \SubstGen{\(M_{\text{new}}\), \(N_{\text{old}}(\varphi)\)}
\(\cup\) \SubstGen{\(M_{\text{old}}\), \(N_{\text{new}}(\varphi)\)}}{
      \ForEach{\(f\mapsto (\upsilon_1, \dots,\upsilon_n)\in N_{\text{old}}(\varphi)\)}{
            \eIf{\TyVars{\((\upsilon_1, \dots, \upsilon_n)\)} \(\subseteq\) \Domain{\(\sigma\)}}{
               \textcolor{ourblueviolet}{
               \uIf{\(|M_{\text{next}}| < \MaxMono\) \label{algo:relative_bound}}{
                  \textcolor{black}{
                  \(M_{\text{next}}(f)\leftarrow M_{\text{next}}(f)\cup\{(\sigma(\upsilon_1),\dots,\sigma(\upsilon_n))\}\)\;
                  \UsedSubst \(\leftarrow \UsedSubst \cup\{\sigma\}\)\;}
               }
               \uElseIf{\(|N_{\text{next}}(\varphi)| \geq \MaxPoly\) }{
                  \(M_{\text{next}} \leftarrow M_{\text{next}}\setminus (M_{\text{old}} \cup M_{\text{new}})\)\;
                  \(N_{\text{next}}(\varphi)\leftarrow N_{\text{next}}(\varphi)\setminus (N_{\text{old}}(\varphi)\cup N_{\text{new}}(\varphi))\)\;
                  \Return \((M_{\text{next}}, N_{\text{next}}(\varphi), \UsedSubst)\)
               }}
            }
            {
               \textcolor{ourblueviolet}{
               \If{\(|N_{\text{next}}(\varphi)| < \MaxPoly\)\label{algo:ref_2}} {
               \textcolor{black}{
               \(N_{\text{next}}(\varphi)(f)\leftarrow N_{\text{next}}(\varphi)(f)\cup\{(\sigma(\upsilon_1),\dots,\sigma(\upsilon_n))\}\)\;
               \UsedSubst \(\leftarrow \UsedSubst \cup\{\sigma\}\)\;
            }}
            }

         }
      }
   }
   \BlankLine

@@

The set of substitutions generated thusly is then iterated over and each substitution is applied to all polymorphic type arguments. The application of a substitution to a polymorphic type argument will only take place if any type variable of the type argument is instantiated by the substitution. This condition filters out all substitution application that would have left the type arguments unchanged.

@@

The total number of type arguments can increase cubicly in the number of type arguments at each iteration and can therefore grow doubly exponentially in the number of iterations. Depending on the shape and size of the input problem and the number of iterations performed, this may be problematic. The limit on the number of new monomorphic type arguments (in \textcolor{ourblueviolet}{blue}) has three components (similar bounds are used for the non-monomorphic type arguments):
\begin{enumerate}
   \item \textcolor{ourblueviolet}{\MonoCap} is an absolute limit on the total number of new type arguments
   \item \textcolor{ourblueviolet}{\MonoMult} is used to allow the total number of type arguments to increase by a certain proportion
   \item \textcolor{ourblueviolet}{\MonoFloor} balances out \MonoMult when the number of type arguments is too low, and guarantees that a minimum number of new type arguments
\end{enumerate}

Additionally, it is possible to either generalise or further refine these bounds. Indeed, the maximum number of newly generated type arguments is fixed per formula, per iteration. This algorithm can easily be modified such that the limit on new type arguments is fixed per iteration, or conversely, be made to limit new type arguments per symbol, per formula, per iteration.

\begin{figure}
\begin{quote}
\begin{algorithm}[H]
\SetKwFunction{SubstGen}{matches}
\Fn(){\SubstGen{\(M, P(\varphi)\)}}{
   \SetKw{And}{and}\SetKw{True}{true}\SetKw{False}{false}
   \SetKwData{SubstLimit}{substitution\_limit}
   \SetKwFunction{Success}{Success}\SetKwFunction{Compatible}{compatible}
   \SetKwFunction{Match}{match}
   \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
   \SetKw{Break}{break}

@@

   \ForEach{\(f\mapsto (\upsilon_1, \dots,\upsilon_n)\in P(\varphi)\)}{
      \ForEach{\( (\tau_1, \dots,\tau_n) \in M(f) \)}{
         \(\sigma\leftarrow \{\}\)\;
         \(i \leftarrow 1\)\;
         \While{\(i \leq n\)}{
            \eIf{\(\Match{\(\upsilon_i\), \(\tau_i\)}\) has the form \Success{\(\sigma_i\)} \And \Compatible{\(\sigma\), \(\sigma_i\)}}{
               \(\strut\sigma\leftarrow\sigma_i\circ\sigma\)
            }
            {\Break}

@@

The various bounds presented here are, to some extent, redundant. For instance, limiting the number of substitutions generated at this step may be sufficient to curb the number of new type arguments, making the related bounds superfluous. Each bound nonetheless has its use cases. For example, sets of problems which lead to few successful matches but many type arguments may benefit from a limit on the number of new type arguments whilst sets of problems where substitution generation is a bottleneck might consider a limit on the number of generated substitution to be more useful.

@@

   \Fn(){\MonoSubst{\(S, V, S_{\text{res}}, \sigma\)}}{

@@

   \eIf{\(V = \emptyset\)}{
    \Return \(S_{\text{res}}\cup\{\sigma\}\)\;%}
   }
   {
      \Let \(\alpha \) \textbf{such that} \(\alpha\in V\)\;
      \ForEach{\(\sigma_{\negvthinspace\Delta}\in S\) \textbf{such that} \(\alpha\in\Domain{\(\sigma_{\negvthinspace\Delta}\)}\) \And \(\Compatible(\sigma,\sigma_{\negvthinspace\Delta})\)}{
         \textcolor{ourblueviolet}{
         \eIf{\(|S_{\text{res}}|<\Limit\)}{
            \textcolor{black}{
            \(S_{\text{res}}\leftarrow \MonoSubst{\(S, V\setminus\Domain{\(\sigma_{\negvthinspace\Delta}\)}, S_{\text{res}}, \sigma_{\negvthinspace\Delta}\circ\sigma\)}\)\;}
         }
         {\Return \(S_{\text{res}}\)}}
      }
      \Return \(S_{\text{res}}\)\;
   }
   }

@@

To monomorphise a polymorphic formula, we first compute its \emph{monomorphising substitutions} (\ref{mono_substs}). A substitution \(\sigma\) is a monomorphising substitution for a formula \(\varphi\) if \(\sigma(\varphi)\) is monomorphic. \ref{mono_substs} is a recursive function, given a set of type variables \(V\) and a pool of substitutions \(S^\text{all}\), it will select a substitution \(\sigma_{\negvthinspace\Delta}\) from \(S^\text{all}\) which instantiates at least one of the type variables in \(V\). It is important that \(\sigma_{\negvthinspace\Delta}\) be compatible with \(\sigma\) so that they may be composed and the function recursively called to instantiate the remaining type variables.

The Zipperposition implementation of this function makes use of a map from type variables to substitutions in place of a simple set to filter the relevant substitutions from \(S^\text{all}\) with greater ease. The \textcolor{ourblueviolet}{max\_substs} exists for two main reasons:
\begin{enumerate}
   \item The iterative monomorphisation algorithm can generate up to \textcolor{ourblueviolet}{max\_substs} new monomorphic clause per initial polymorphic clause. Generating an excessive number of new clauses can lead to the downstream prover being flooded.
   \item This function is the most explosive part of the algorithm. If \(S^\text{all}\) contains \(n\) substitutions which each instantiate exactly one of \(v\) different type variable, up to \(n^v\) monomorphising substititions may be generated. This should be considered in light of the fact that the total number of type variables (used to generate these substitutions) is already doubly exponential in the number of loop iterations.
\end{enumerate}

@@

Once the newly generated monomorphic formulae are output, an extra post-processing step may be required. Indeed, monomorphic provers, such as E do not necessarily support type constructors. Consequently, transforming the monomorphic formulae into equivalent ones without type constructors may be necessary as it was for the Zipperposition implementation. This transformation is largely syntactic, it is documented in [insert names] 's paper [insert citation].

>>>>

\textbf{ TODO rename non-monomorphic maps from \(P\) to \(N\) in high-level description section}

@@

The data structures used in the pseudocode are based on the ones used in the high-level description. Instead of \(M\) a map from symbols to monomorphic type argument tuples, we have \(M_\text{old}\) and \(M_\text{new}\) which play the same role whilst also distinguishing between type arguments having already been matched against and those that have not. Similarly, \(N_\text{old}\) and \(N_\text{new}\) replace \(N\) the map from formulae to symbols to non-monomorphic type argument tuples. Finally, a set \(S\) of the substitutions generated by the matchings is kept track of throughout the iterations, until it is used in the final phase.

All sets referenced in the pseudocode are assumed to be finite. Additionally, this algorithm relies on ``primitives'' whose implementation depend on the specifics of the grammar and logic used. They will therefore not be expanded upon. Functions computing the following are assumed to be available: 
\begin{enumerate}
   \item[\labelitemi] extracting the initial type argument maps from the input problem
   \item[\labelitemi] gathering type variables from a type tuple
   \item[\labelitemi] type matching
   \item[\labelitemi] returning the type variables instantiated by a substitution
   \item[\labelitemi] testing the compatibility of two substitutions
   \item[\labelitemi] substitution composition
   \item[\labelitemi] substitution application to formulae
\end{enumerate}

@@

\SetKwFunction{GenFormulae}{generate\_mono\_formulae}
\SetKw{ST}{such that}

\begin{figure}
\begin{quote}
\begin{algorithm}[H]
\Fn(){\(\GenFormulae(\Phi, S)\)}{
   \SetKw{And}{and}\SetKw{True}{true}\SetKw{False}{false}\SetKw{To}{to}

@@

   \ForEach{\(\varphi\in\Phi\) \ST $\varphi$ is non-monomorphic}{
      \ForEach{\(\sigma\in \MonoSubst(\AllSubst, \TyVars{\(\varphi\)}, \emptyset, \{\})\)}{
         \textcolor{ourblueviolet}{
         \eIf{\(|\NewFormulae|<\Limit\)}{
            \textcolor{black}{
            \(\NewFormulae\leftarrow\NewFormulae\cup \{\sigma(\varphi)\}\)\;}
         }
         {\Return \NewFormulae}
      }}
   }

@@

   \Return \NewFormulae
}
\end{algorithm}
\end{quote}
\caption{Pseudocode for monomorphic formula generation}
\label{gen_formulae}
\end{figure}

The iterative monomorphisation algorithm \ref{iter_mono}, as implemented in Zipperposition, has two main phases each consisting of a separate loop. The first loop applies a \emph{monomorphisation step} to each formula \ref{mono_step} for each iteration of the outermost for loop until the user-set limit it reached. This bound is the only limit which is necessary for the algorithm to terminate. At the end of each of these iterations, the old and new type argument maps are updated with newly generated types.
%Updating the maps progressively over the course of an iteration would be feasible but would result in an uneven treatment of formulae. Type arguments in a formula processed later could be matched against monomorphic type arguments generated at an earlier time in the same iteration. As a result, the later a formula is processed the more new type arguments it could generate.

Once the first phase is completed, the substitutions used to create new type arguments are combined to form \emph{monomorphising substitutions} (\ref{gen_formulae}). These substitutions are subsequently applied to the input problem's polymorphic formulae. The possible number of new (monomorphic) formulae resulting from this process is high but can be limited by setting a maximal number of new formulae. In the Zipperposition implementation, this bound is one of the more important ones as the performance of the E prover can be significantly affected by the number of clauses it is given.

\SetKwData{MonoCap}{mono\_cap}\SetKwData{MonoMult}{mono\_mult}\SetKwData{MonoFloor}{mono\_floor}
\SetKwData{PolyCap}{poly\_cap}\SetKwData{PolyMult}{poly\_mult}\SetKwData{PolyFloor}{poly\_floor}
\SetKwData{TypeVars}{type\_variables}
\SetKwData{MaxMono}{max\_mono\_args}
\SetKwData{MaxPoly}{max\_poly\_args}
\SetKwData{UsedSubst}{\(S\)}

@@

\SetKwFunction{Max}{max}\SetKwFunction{Min}{min}
\SetKwFunction{Domain}{subst\_domain}\SetKwFunction{Len}{len}
\SetKwFunction{SubstGen}{matches}\SetKwFunction{TyVars}{type\_variables}
\SetKwFunction{FMonoStep}{formula\_mono\_step}

   \Fn(){\(\FMonoStep(\varphi, M_{\text{old}}, M_{\text{new}}, N_{\text{old}}(\varphi), N_{\text{new}}(\varphi)\)}{

@@

   \(\TypeVars \leftarrow \TyVars(\varphi)\)\;

   \textcolor{ourblueviolet}{
   \(\MaxMono \leftarrow \Max(\Min(\MonoFloor, |M_{\text{old}}\cup M_{\text{new}}| \cdot \MonoMult), \MonoCap)\)\;
   \(\MaxPoly \leftarrow \Max(\Min(\PolyFloor, |N_{\text{old}}(\varphi)\cup N_{\text{new}}(\varphi)| \cdot \PolyMult), \PolyCap)\)\;
   }

@@

   \ForEach{\(\sigma\in \SubstGen(M_{\text{new}}, N_{\text{new}}(\varphi))
   \cup \SubstGen(M_{\text{new}}, N_{\text{old}}(\varphi))
   \cup \SubstGen(M_{\text{old}}, N_{\text{new}}(\varphi))\)}{
      \ForEach{\(f\mapsto (\upsilon_1, \dots,\upsilon_n)\in N_{\text{old}}(\varphi)\)}{
            \eIf{\(\TyVars((\upsilon_1, \dots, \upsilon_n)) \subseteq \Domain(\sigma)\)}{
               \textcolor{ourblueviolet}{
               \uIf{\(|M_{\text{next}}| < \MaxMono\)}{
                  \textcolor{black}{
                  \(M_{\text{next}}(f)\leftarrow M_{\text{next}}(f)\cup\{(\sigma(\upsilon_1),\dots,\sigma(\upsilon_n))\}\)\;
                  \UsedSubst \(\leftarrow \UsedSubst \cup\{\sigma\}\)\;}
               }
               \uElseIf{\(|N_{\text{next}}(\varphi)| \geq \MaxPoly\) }{
                  \(M_{\text{next}} \leftarrow M_{\text{next}}\setminus (M_{\text{old}} \cup M_{\text{new}})\)\;
                  \(N_{\text{next}}(\varphi)\leftarrow N_{\text{next}}(\varphi)\setminus (N_{\text{old}}(\varphi)\cup N_{\text{new}}(\varphi))\)\;
                  \Return \((M_{\text{next}}, N_{\text{next}}(\varphi), \UsedSubst)\)
               }}
            }{\textcolor{ourblueviolet}{
               \If{\(|N_{\text{next}}(\varphi)| < \MaxPoly\)}{ \textcolor{black} {
               \(N_{\text{next}}(\varphi)(f)\leftarrow N_{\text{next}}(\varphi)(f)\cup\{(\sigma(\upsilon_1),\dots,\sigma(\upsilon_n))\}\)\;
               \(\UsedSubst \leftarrow \UsedSubst \cup\{\sigma\}\)\;
               }}
            }}
      }
   }

@@

The set of substitutions generated thusly is then iterated over and each substitution is applied to all polymorphic type arguments. The application of a substitution to a tuple of polymorphic type argument only takes place if least one type variable of a type in the tuple is instantiated by the substitution. This condition filters out all substitution application that would have otherwise left the type arguments unchanged.

@@

The total number of type arguments can increase cubicly in the number of type arguments at each iteration and can therefore grow doubly exponentially in the number of iterations. Depending on the shape and size of the input problem and the number of iterations performed, this may be problematic. Introducing bounds can help solve this potential issue. The limit on the number of new monomorphic type arguments (in \textcolor{ourblueviolet}{blue}) has three components (similar bounds are used for the non-monomorphic type arguments):
\begin{enumerate}
   \item \textcolor{ourblueviolet}{\MonoCap} is an absolute limit on the total number of new type arguments
   \item \textcolor{ourblueviolet}{\MonoMult} is used to allow the total number of type arguments to increase by a certain proportion
   \item \textcolor{ourblueviolet}{\MonoFloor} balances out \MonoMult when the number of type arguments is too low, and guarantees that a minimum number of new type arguments
\end{enumerate}

Additionally, it is possible to either generalise or further refine these bounds. Indeed, the maximum number of newly generated type arguments is fixed per formula, per iteration. This algorithm can easily be modified such that the limit on new type arguments is fixed per iteration. Conversely, it can be made to limit new type arguments per symbol, per formula, per iteration.

\begin{figure}
\begin{quote}
\begin{algorithm}[H]
\SetKwFunction{SubstGen}{matches}
\Fn(){\(\SubstGen(M, P(\varphi))\)}{
   \SetKw{And}{and}\SetKw{True}{true}\SetKw{False}{false}
   \SetKwData{SubstLimit}{substitution\_limit}
   \SetKwFunction{Success}{Success}\SetKwFunction{Compatible}{compatible}
   \SetKwFunction{Match}{match}
   \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
   \SetKw{Break}{break}

@@

   \ForEach{\(f\mapsto (\upsilon_1, \dots,\upsilon_n)\in P(\varphi)\)}{
      \ForEach{\( (\tau_1, \dots,\tau_n) \in M(f) \)}{
         \(\sigma\leftarrow \{\}\)\;
         \(i \leftarrow 1\)\;
         \While{\(i \leq n\)}{
            \eIf{\(\Match(\upsilon_i, \tau_i)\) has the form \(\Success(\sigma_i)\) \And \(\Compatible(\sigma, \sigma_i)\)}{
               \(\strut\sigma\leftarrow\sigma_i\circ\sigma\)
            }
            {\Break}

@@

The various bounds presented here are, to some extent, redundant. For instance, limiting the number of substitutions generated at this step may be sufficient to curb the number of new type arguments, making the related bounds superfluous. Each bound nonetheless has its use cases. For example, sets of problems which lead to few successful matches but many type arguments may benefit from a limit on the number of new type arguments whilst sets of problems where substitution generation is more explosive might consider a limit on the number of generated substitution to be more useful.

@@

   \Fn(){\(\MonoSubst(S, V, S_{\text{res}}, \sigma)\)}{

@@

   \eIf{\(V = \emptyset\)}{
    \Return \(S_{\text{res}}\cup\{\sigma\}\)\;%}
   }
   {
      \Let \(\alpha \) \ST \(\alpha\in V\)\;
      \ForEach{\(\sigma_{\negvthinspace\Delta}\in S\) \ST \(\alpha\in\Domain(\sigma_{\negvthinspace\Delta})\) \And \(\Compatible(\sigma,\sigma_{\negvthinspace\Delta})\)}{
         \textcolor{ourblueviolet}{
         \eIf{\(|S_{\text{res}}|<\Limit\)}{
            \textcolor{black}{
            \(S_{\text{res}}\leftarrow \MonoSubst(S, V\setminus\Domain(\sigma_{\negvthinspace\Delta}), S_{\text{res}}, \sigma_{\negvthinspace\Delta}\circ\sigma)\)\;}
         }
         {\Return \(S_{\text{res}}\)}}
      }
      \Return \(S_{\text{res}}\)\;
   }
   }

@@

To monomorphise a polymorphic formula, we first compute its \emph{monomorphising substitutions} (\ref{mono_substs}). A substitution \(\sigma\) is a monomorphising substitution for a formula \(\varphi\) if \(\sigma(\varphi)\) is monomorphic. It is a recursive function, given a set of type variables \(V\) and a pool of substitutions \(S^\text{all}\), it will select a substitution \(\sigma_{\negvthinspace\Delta}\) from \(S^\text{all}\) which instantiates at least one of the type variables in \(V\). It is important that \(\sigma_{\negvthinspace\Delta}\) be compatible with \(\sigma\) so that they may be composed and the function recursively called to instantiate the remaining type variables.

The Zipperposition implementation of this function makes use of a map from type variables to substitutions in place of a set to filter the relevant substitutions from \(S^\text{all}\) more efficiently. The \textcolor{ourblueviolet}{max\_substs} exists for two main reasons:
\begin{enumerate}
   \item The iterative monomorphisation algorithm can generate up to \textcolor{ourblueviolet}{max\_substs} new monomorphic clause per initial polymorphic clause. Generating an excessive number of new clauses can lead to the downstream prover being flooded.
   \item This function is the most explosive part of the algorithm. If \(S\) contains \(n\) substitutions which each instantiate exactly one of \(v\) different type variable, up to \(n^v\) monomorphising substititions may be generated. This should be considered in light of the fact that the total number of type argument tuples, which  re used to generate \(S\), is doubly exponential in the number of loop iterations.
\end{enumerate}

@@

Once the newly generated monomorphic formulae are output, an extra post-processing step may be required. Indeed, monomorphic provers, such as E, do not necessarily support type constructors. Consequently, transforming the monomorphic formulae into equivalent ones without type constructors may be necessary as it was for the Zipperposition implementation. This transformation is largely syntactic and relatively simple. It is documented in [insert names] 's paper [insert citation, and section because it isn't the main topic].
